04_analysis

### 2023-05-25 04_analysis.Rmd
more gam and glm varieties

Required Packages
```{r}
library(tidyverse)
```

@@

## ignore??
```{r drift-8-histogram}
ggplot(data = pivot3, aes(Hz, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 1, color = "black") +
  theme_classic() +
  ggtitle("Distribution of Third-Octave Level (TOL) Intensities: Buoy 8") +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_manual(name = "Beaked Whale Presence", 
                    labels = c("0" = "Absent", "1" = "Present"),
                    values = c("0" = "bisque1",
                               "1" = "brown4")) +
  theme(legend.position = c(0.6,0.7)) +
  facet_wrap(~TOL, labeller = as_labeller(TOL.labs))
```

@@

IS THIS HELPFUL??
Whale Occurrences
```{r}
# Filter Drift 8 by Whale presence and absence
W1 <- tracks_SS_08 %>%
  filter(BWpresence == 1)
W0 <- tracks_SS_08 %>%
  filter(BWpresence == 0)

@@

# Salinity at 400 m
ggplot(data = allDrifts, aes(sal400, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 0.005, color = "black")
```

>>>>

Required Packages
```{r}
library(tidyverse)
library(mgcv)
```

@@

## ignore??
```{r drift-8-histogram}
#ggplot(data = pivot3, aes(Hz, fill = factor(BWpresence))) +
#  geom_histogram(binwidth = 1, color = "black") +
#  theme_classic() +
#  ggtitle("Distribution of Third-Octave Level (TOL) Intensities: Buoy 8") +
#  scale_x_continuous(name = "Sound Intensity (dB)") +
#  scale_fill_manual(name = "Beaked Whale Presence", 
#                    labels = c("0" = "Absent", "1" = "Present"),
#                    values = c("0" = "bisque1",
#                               "1" = "brown4")) +
#  theme(legend.position = c(0.6,0.7)) +
#  facet_wrap(~TOL, labeller = as_labeller(TOL.labs))
```

@@

IS THIS HELPFUL??
Whale Occurrences
```{r}
# Filter Drift 8 by Whale presence and absence
W1 <- tracks_SS_08 %>%
  filter(Wpresence == 1)
W0 <- tracks_SS_08 %>%
  filter(Wpresence == 0)

@@

# Salinity at 400 m
ggplot(data = allDrifts, aes(sal400, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 0.005, color = "black")
```
### GAMs

```{r}
model3 <- gam(BWpresence ~ depth + TOL_125 + dist2shore + mldTemp, 
              data = allDrifts, family = binomial)
plot(model3)

#try fitting linear model

lm_depth <- lm(BWpresence ~ depth, data = allDrifts)
ggplot(allDrifts, aes(sst_mean, chlorophyll_mean)) +
  geom_point() +
  geom_smooth()
```
Categorical term
```{r}
# ALL variables, GAM
model1 <- gam(BWpresence>0 ~ s(BB_20.24000) + s(TOL_63) + s(TOL_125) + s(TOL_2000)
              + s(TOL_5000) + s(TOL_20000) + s(depth) + s(mldDepth) + s(Latitude)
              + s(Longitude) + s(sst_mean) + s(curl_mean) + s(dist2slope) 
              + s(bathy_slope) + s(dist2shore) + s(chlorophyll_mean) + s(ssh_mean) 
              + s(mldTemp) + s(ttDepth) + s(ttTemp) + s(temp400) + s(sal400),
              data = allDrifts, family = binomial, method = "REML")

# ALL variables, GLM
model1GLM <- glm(BWpresence ~ BB_20.24000 + TOL_63 + TOL_125 + TOL_2000
              + TOL_5000 + TOL_20000 + depth + mldDepth + Latitude
              + Longitude + sst_mean + curl_mean + dist2slope
              + bathy_slope + dist2shore + chlorophyll_mean + ssh_mean 
              + mldTemp + ttDepth + ttTemp + temp400 + sal400,
              data = allDrifts, family = binomial)

# By species? doesnt work, may need to create dummy variables
model2 <- gam(BWpresence ~ s(sst_mean, by = as.factor(species)), 
              data = allDrifts, family = binomial, method = "REML")

# Soundscape metrics only, GAM
model4 <- gam(BWpresence ~ s(BB_20.24000, k = 10) + 
                s(TOL_63, k = 10) + s(TOL_125, k = 10) + 
                s(TOL_2000, k = 10) + s(TOL_5000, k = 10) 
              + s(TOL_20000, k = 10),
              data = allDrifts, family = binomial, method = "REML")

# Soundscape metrics only, GLM
model4GLM <- glm(BWpresence ~ BB_20.24000 + TOL_63 + TOL_125 + TOL_2000
              + TOL_5000 + TOL_20000,
              data = allDrifts, family = binomial)

model125 <- gam(BWpresence ~ s(TOL_125, k = 10),
              data = allDrifts, family = binomial, method = "REML")
plot(model125, trans = plogis)
summary(model125)
gam.check(model125)

# Environmental var only
model5 <- gam(BWpresence ~ s(depth) + s(mldDepth) + s(Latitude)
              + s(Longitude) + s(sst_mean) + s(curl_mean) + s(dist2slope) 
              + s(bathy_slope) + s(dist2shore) + s(chlorophyll_mean) + s(ssh_mean) 
              + s(mldTemp) + s(ttDepth) + s(ttTemp) + s(temp400) + s(sal400),
              data = allDrifts, family = binomial, method = "REML")

plogis(coef(model1[1]))
summary(model1)
summary(model4)
summary(model5)
summary(model4GLM)
plot(model4GLM)
gam.check(model1)
gam.check(model4)

#Log odds scale (when used with family = binomial)
plot(model1, pages = 6)
plot(model4, pages = 4, trans = plogis)

#Probability scale + centered on the intercept + adding intercept-related uncertainty with seWithMean arg
plot(model1, pages = 14, trans = plogis,
     shift = coef(model1)[1])#,
     #seWithMean = TRUE)

plot(model2, pages = 1, trans = plogis,
     shift = coef(model2)[1])

plot(model4, pages = 6, trans = plogis,
     shift = coef(model4)[1])
```
